Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motions Characteristic to Vertical Axis Turbines

### 2016-03-01 paper.tex
Wrapping and whitespace

\begin{abstract}Vertical Axis Wind Turbines (VAWTs) have been used for extracting of energy either
    from sea or wind. A simpler design, low cost of maintenance and the
    capability of working independent from the direction of the flow are some of
    the most important advantages over conventional Horizontal Axis Wind Turbines
    (HAWTs). Complex and unsteady fluid dynamics are not easy to model for
    simulation tools. One of the most relevant phenomena is the dynamic stall
    which is inherently to unsteady conditions, this work is focused in its
    study. Dynamic stall is usually used as a passive control for VAWT operating
    conditions, hence the importance of a good understanding. In this study,
    \textbf{a coupled model is implemented with the open-source CFD toolbox
    OpenFOAM for solving the governing equations; an actuator line model is used
    for solving velocity flow field and a dynamic stall model for force
    interactions.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as
    H-rotor type. Numerical results show reasonable agreement with experimental data
	for pitching motion. In general, simulations allow to characterize dynamic stall effects 
	and their magnitude.

>>>>

\begin{abstract}

    Vertical Axis Wind Turbines (VAWTs) have been used for extracting of energy
    either from sea or wind. A simpler design, low cost of maintenance and the
    capability of working independent from the direction of the flow are some of
    the most important advantages over conventional Horizontal Axis Wind
    Turbines (HAWTs). Complex and unsteady fluid dynamics are not easy to model
    for simulation tools. One of the most relevant phenomena is the dynamic
    stall which is inherently to unsteady conditions, this work is focused in
    its study. Dynamic stall is usually used as a passive control for VAWT
    operating conditions, hence the importance of a good understanding. In this
    study, \textbf{a coupled model is implemented with the open-source CFD
    toolbox OpenFOAM for solving the governing equations; an actuator line model
    is used for solving velocity flow field and a dynamic stall model for force
    interactions.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as
    H-rotor type. Numerical results show reasonable agreement with experimental
    data for pitching motion. In general, simulations allow to characterize
    dynamic stall effects  and their magnitude.
